---
title: "glicko_vs_spread"
output: html_document
---

```{r}
library(tidyverse)

# Umm...how did I make this NFL results csv?
results <- read_csv('../../py-glicko/examples/nfl/nfl_results.csv')
spreads <- read_csv('../../EndGame/py-endgame/nfl_spreads.csv')

against_spread <- results %>%
  inner_join(spreads, by = c('team' = 'home', 'opponent' = 'away', 'week', 'season')) %>% 
  mutate(home_mov = team_score - opponent_score, favored_pct = win_prob - .5)
```

```{r}
train_season <- 2012
prob_to_spread <- lm(spread ~ poly(favored_pct, 3), data = against_spread %>% filter(season < train_season))
summary(prob_to_spread)
```


```{r}
win_probs <- seq(-.5, .5, by = .01)
ggplot(tibble(win_probs, movs = -predict(prob_to_spread, newdata = tibble(favored_pct = win_probs)))) +
  geom_line(aes(x = win_probs, y = movs)) +
  labs(x = "win prob - 0.5", y = "Predicted home MOV")
```

```{r}
against_spread <- against_spread %>%
  mutate(predicted_home_mov = -predict(prob_to_spread, newdata = tibble(favored_pct = favored_pct))) %>% 
  mutate(residual = home_mov - predicted_home_mov) %>% 
  # Spread is the number of points given to (or subtracted from if negative) the home team
  mutate(home_beat_spread = team_score + spread > opponent_score) %>% 
  mutate(i_pick_home = predicted_home_mov > -spread) %>% 
  mutate(beat_spread = !xor(home_beat_spread, i_pick_home))
```

Does this get better vs the spread over time? Or deeper into seasons? Eh, not really...

```{r}
against_spread %>% group_by(season) %>% summarise(beat_spread_pct = mean(beat_spread)) %>% 
  ggplot() +
  geom_line(aes(x = season, y = beat_spread_pct)) +
  labs(title = "Do I get better vs. spread over time?", y = "Win % over spread")

against_spread %>% filter(season >= train_season) %>%
  group_by(week) %>%
  summarise(vs_spread = mean(beat_spread)) %>%
  ggplot() +
  geom_line(aes(x = week, y = vs_spread)) +
  ylim(c(0, 1)) +
  labs(title = "vs. spread further into the season?", y = "Win % over spread")
```

Overall win %:

```{r}
against_spread %>% filter(season >= train_season) %>% .[['beat_spread']] %>% mean()
```

# Explore

Check some different views. None are very interesting except the last one.
This beats the spread fairly frequently in close games, loses in blowouts.
That's interesting b/c other facets don't show similar problems - 
games with predicted large MOVs (by the spread, or by my model)
are still roughly even vs. the spread, it's just games that ended up being blowouts

```{r}
against_spread %>% ggplot() + geom_point(aes(x = spread, y = predicted_home_mov, col = beat_spread))
against_spread %>%
  ggplot() + geom_density(aes(x = predicted_home_mov, col = beat_spread))
against_spread %>%
  ggplot() + geom_density(aes(x = spread, col = beat_spread))
against_spread %>%
  filter(season >= train_season) %>%
  ggplot() + geom_density(aes(x = home_mov, col = beat_spread))
```
